\vspace{-5em}\Huge Generation of a Synthetic Population and Travel Network for an Agent Based Traffic Simulation

### 2023-09-25 paper/paper.tex
Small grammar updates

    \quad Fortunately, nowadays, traffic simulation has started to move in the direction of agent based models \cite{ile-de-france}. These models function on a lower level, looking at the activities of people and their interactions \cite{agent-based-model}. However, in order to compute such specific predictions about a population, a lot of input data is required. Similar to any form of a simulation, you can't create information, just analyze it. Thus, the accuracy of the traffic demand determines the accuracy of the output. The best way to go about solving this problem is by sampling large populations down to an accurate, but still computable size. Still, very few governments openly release sufficient data about their populations for traffic modeling. One example, can be seen in France, where an open-source pipeline was created for this exact scenario of sampling \cite{ile-de-france}. The pipeline uses stochastic\footnote{Use of random probability to determine a quantity} models, such as the Monte Carlo simulation to sample an even size of the data of France. In their case, the model results in data that is just 5\% of the size of Île-de-France's population. Having such a tool allows researchers to analyze existing traffic flows through agent based simulations, such as MATSIM\footnote{Open source framework for implementing agent based transport simulations} \cite{matsim}.\\

@@

    \section{Theory}
    \quad As stated previously, the generation of traffic demand depends on a couple of very crucial background topics. Zoning policies are very important, as they are the center of how many cities around the world are designed. This program allows for Euclidean zoning types with both hierarchical and flat zoning. The specifics of what these mean will be analyzed later, but in general both policies are very rigid. In fact, various papers link these types of zoning policies to urban sprawl \cite{zoning}. Though various other forms of zoning policies exist, euclidean zoning is by far the most popular and easy to implement.\\

@@

    \quad The two other major forms of zoning that are newer and harder to enforce are performance based and form based zoning policies. Performance based zoning is more of a points based system that determines the criteria for a building existing within a zone as codes of compliance \cite{zoning}. These could include facilities, such as public restrooms or other standards for environmental reasons. With such a diverse and complex structure, this policy struggles to be utilized widely. Form based zoning on the other hand, is new and on the rise. This policy enforces a building's compliance through dimensions and its size \cite{zoning}. \\

@@

    \quad Cellular automaton was an idea thrown around to allow zoning cells to connect with one another. The pros of this idea is that it was a way to traverse the grid and set different points to a zone if not in use. As a whole, cellular automaton takes in a grid of cells and assigns values, such as on or off to these cells. Starting at a cell, the program finds its neighbors and applies a function to them to determine if their value should be changed. Now, this algorithm is actually very popular for binary systems, such as Conway's Game of Life: seen in Figure \ref{fig:cellular-automaton}. However, it tends to get very complicated very quickly when applied to a system on a higher order. Especially, a system such as configurable zoning, which does not have a finite quantity to it. If a user added more zones to the city, the program would struggle to make this functionality possible. Therefore, cellular automaton was finally determined not applicable to our zoning scenario.\\

@@

    \quad Unlike zoning, making a network is actually much simpler. The two main steps comprise of creating nodes within the existing zones and making a road system that connects them together. These roads need to ensure that a person can travel from any given node to another. In our case, a node is simply a building. To determine the density of nodes in one district, several values are provided from the user's configurations. These include the average size of one building in the zone. However, this quantity is mainly used to produce the size of the zones. As far as the network goes, the generation of nodes strictly relies on the area of a building in any subzone and the area of a building in the top level zone. If no subzones are present, the types of buildings in the district will be a list of one type (the top level zone). After determining the types of buildings from the configuration file, the program randomly picks a building type and subtracts the area of it from the district's remaining area until no more buildings can be added. However, one important note is that the program also ensures that atleast one building of the top level zone's type is present. This is done by scanning neighbors, as seen in section \ref{zoning}. Once the nodes are picked out, they are converted to type Location\footnote{A class defined to make referencing node attributes easier} in the code.\\

@@

    \quad Figure \ref{fig:first-zoning} illustrates the first zoning implementation of the program. This was done by taking the grid and randomly drawing a rectangle by selecting two points. These points were the top left and bottom right corners of the shape. Each rectangle was also constrained to a maximum area. This prevented the rectangle from taking over the entire plot of land. After the rectangle was created, the program selected the largest remaining plot of land. Each plot of land was defined as one edge of the existing rectangle to the end of the city. Repeating this process with the new plot of land allowed for the program to consistently draw out equal sized portions of zones. However, there were several problems with this algorithm. First off, it was overly complex. Looking for available plot of land was computer intensive and was not always guaranteed to be large enough. In addition, the rectangles did well for three zones, but quickly deteriorated after adding more.

@@

    In this formula, A represents the area of a district, B is the average size of a building in a given zone, and M is the minimum number of buildings within a district. After computing A for every zone, the algorithm picked out the minimum area of a district and set the cell length to \(\sqrt{A}\). This method worked fairly well, but still had several issues. For one, the zones were all of the same size blocks. Also, apart from the predominant zone, all the other districts were usually one block large. In most modern cities, the size of a district depends on its zone, so this was definitely a problem to solve.

@@

    \[C = A \div m\]

    In this formula, C represents the maximum number of cells in a given district, A is once again the district area determined above, and m is the minimum building size of all given zones.

@@

    \[N = A \div B\]

    In this formula, N is the number of nodes, A is the area of a district\footnote{The district in this iteration was just a rectangle}, and B is the area of one average building in the zone. Thus, each zone had a different density of dots within it.

    \begin{figure}[H]
        \centering
        \vspace{-1em}
        \includegraphics[width=0.45\textwidth]{images/secondzoningwnodes.png}
        \caption{Second iteration of network without roads}
        \label{fig:second-nodes}
    \end{figure}

    \begin{figure}[H]
        \centering
        \vspace{-1em}
        \includegraphics[width=0.45\textwidth]{images/secondzoningwnetwork.png}
        \caption{Second iteration of complete network with roads}
        \label{fig:second-network}
    \end{figure}

    \quad As seen in Figure \ref{fig:second-nodes} the distribution of buildings stayed relatively similar to the previous iteration. The only difference was in the display of them to the user. The larger dots in the previous iteration made the detail of their location less precise. Thus, a reduction of size was a simple fix to make the data more clear. In Figure \ref{fig:second-network} it can be seen that roads are finally implemented. However, this is not very representative of a real city's road system. The reason for this is because of the way roads are assigned. In each district, one central hub is chosen, and every other node is connected to that hub with a district road. These district roads were not configurable and only allowed cars to travel on them. In addition, the mess of white lines is just a very inefficient highway system. In this case, the central hubs are shuffled and connected to the next hub in a list. Despite the ineffective nature of the road network, it still functioned to complete one of the fundamental goals: having a system that allowed a person to move from one building to another. This data was ported onto MATSIM\cite{matsim}, but it did not reveal anything surprising or interesting. \\

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.45\textwidth]{images/finalzoningwnodes.png}
        \caption{Final iteration of the nodes}
        \label{fig:final-zoning-nodes}
    \end{figure}

@@

    \quad The final iteration of the road system was much neater and more organized with the help of object oriented programming. The nodes were finally calculated with a different algorithm. Given a district, with a size determined in Section \ref{zoning-implementation}, the land was iterated over to find the placement of zones. The program continuously picked random points in the districts and subtracted the area of the placed building from the remaining area in the district. For a zone with subzones within it, the program simply chose the building type from a list of all possible buildings. Once, again all of this information was configurable by the user. In addition, the program placed atleast one building from the top level zone in this case by checking neighbors within the district for a building of the correct type. If no building was found, then the program inserted a top level building automatically. This resulted in the variation of dot colors as seen in the plot. The different color represent different building types. Finally, this iteration also added our better road network. This was heavily described in section \ref{network}. However, to put it simply the baseline road network is just a minimum spanning tree that connects nodes in a district. The highways are a minimum spanning tree that connect a number of central hubs in each district, defined in the configuration file. Each type of road also has attributes of speed, capacity, and number of lanes, which can be changed by configurations as well.

@@

    In this equation, R represents the number of additional roads, r is the number of existing roads in a district or highway network, and M is the meshing ratio defined by the configuration file. Once the number of additional roads is determined, the algorithm goes through all edges that have not been added to the minimum spanning tree from the complete graph\footnote{A graph with all possible edges} and randomly chooses roads to add to the network till the R value has been satisfied.

@@

    \quad The final step of the traffic demand generation is creating the people or agents. These agents require schedules to fulfill the requirements of MATSIM\cite{matsim}. Each schedule must be realistic and also highly configurable. To do this, the program takes in inputs of timings of each activity in particular zones. Thus, when nodes are generated and converted to Location objects, as seen in Section \ref{network}, they also are randomly assigned values of times for when agents use the facilities. In addition, the configuration file also determines the number of workers within each zone. Finally, the configuration file also determines the percentage of people that work from within their district and the percentage of people that work from home. These pieces of data are essentially sifted through with if statements to determine the timings of a person's job. However, to first assign an agent a location to work is a little more complicated. First, for each household, all jobs nearby are sorted from closest to furthest. Then, each location is assigned a weight based on its proximity or index. Finally, the algorithm chooses a random job with the weights assigned. A person is determined to work at home by checking if a random quantity in the range \([0, 1)\) is less than the ratio of the population that work from home. If so, the step of picking a job is skipped. Finally, agents are put through an infinite loop, which has a third of a chance to end every iteration. This loop adds leisure activities after work that fit into the schedule of the agent. These activities are chosen the same way as work. However, this time the number of workers working at the facility does not matter. To increase diversity in schedules, the variability of a person arriving to and leaving from a location can also be modified. Since buildings can have multiple types, people can work and leisure at the same location. This allows for a completely customizable schedule of times and places.

@@

    \quad Once the traffic demand has been generated, the information can be utilized to simulate the city. As mentioned prior, MATSIM \cite{matsim} is the agent based model of choice for this program. The data produced is converted into XML format for the simulation. These files are stored within the output/ directory and can be used by MATSIM later on. In fact, the code even automatically executes the simulation if the executable file is found in the main directory. These results are stored in the new outputs/ directory and can be analyzed for research. In fact, most of this information can be fed into a wrapper script for implementation of a machine learning algorithm. Such an algorithm can stochastically change the configurations and find the effects on traffic. Simultaneous Perturbation Stochastic Approximation is an example of such an algorithm to find a more ideal city from a starting slate. However, with such a large set of configuration options, a Simulated Annealing algorithm may work better. Despite this, the data from a random city can be found in the figures below.

@@

    \quad The leg histogram found in Figure \ref{fig:leg-histogram} is one of an inefficient city. Nonetheless, it enunciates the different values and makes it clear the meaning of the colors. The red lines represent agents departing from their source locations for any new task in the schedule. The green lines are the number of agents on route. Finally, the blue lines are the number of agents arriving at their destinations. Since this simulation was of a configuration that restricted the bike's speed to 0.5 meter per second, the people took a long time to reach their destinations. It is also clear that due to the few number of jobs in this city, many agents are leaving at around the same time to get to work, however, the schedules in the evenings get a lot more sparse. \\

@@

    \quad This project was successful in that it captured and modeled a theoretical city from a very easily configurable json file. Of course, there are always things that can be done better. For one, the zoning policy is enforced on pixels that are the size of the largest possible buildings. However, this doesn't work in all cases. Sometimes, housing zones may take up less space than one industrial building. To account for this, a better system of DFS needs to be implemented to ensure that each zone can break down to different size cells. In addition, one of the major flaws of this program is multi-modal transportation. Though, the pipeline does take biking and cars into account, a network of public transport systems would be an amazing update for traffic engineers. The only difficulty with this is allowing the user to easily configure the timings of each public transport line. Despite these setbacks, the project went through several iterations and does a great job of capturing some of the most complex details of a city. Along with fitting in smoothly with MATSIM, the project can be very much added upon with a wrapper stochastic approximation script. This could reveal aspects of city design that may not be known about yet.

>>>>

    \quad Fortunately, nowadays, traffic simulation has started to move in the direction of agent based models \cite{ile-de-france}. These models function on a lower level, looking at the activities of people and their interactions \cite{agent-based-model}. However, in order to compute such specific predictions about a population, a lot of input data is required. Similar to any form of a simulation, you can't create information, just analyze it. Thus, the accuracy of the traffic demand determines the accuracy of the output. The best way to go about solving this problem is by sampling large populations down to an accurate, but still computable size. Still, very few governments openly release sufficient data about their populations for traffic modeling. One example can be seen in France, where an open-source pipeline was created for this exact scenario of sampling \cite{ile-de-france}. The pipeline uses stochastic\footnote{Use of random probability to determine a quantity} models, such as the Monte Carlo simulation to sample an even size of the data of France. In their case, the model results in data that is just 5\% of the size of Île-de-France's population. Having such a tool allows researchers to analyze existing traffic flows through agent based simulations, such as MATSIM\footnote{Open source framework for implementing agent based transport simulations}\cite{matsim}.\\

@@

    \section{Theory}
    \quad As stated previously, the generation of traffic demand depends on a couple of very crucial background topics. Zoning policies are very important, as they are the center of how many cities around the world are designed. This program allows for Euclidean zoning types with both hierarchical and flat zoning. The specifics of what these mean will be analyzed later, but in general both policies are very rigid. In fact, various papers link these types of zoning policies to urban sprawl \cite{zoning}. Though various other forms of zoning policies exist, Euclidean zoning is by far the most popular and easy to implement.\\

@@

    \quad The two other major forms of zoning that are newer and harder to enforce are performance-based and form-based zoning policies. Performance-based zoning is more of a points based system that determines the criteria for a building existing within a zone as codes of compliance \cite{zoning}. These could include facilities, such as public restrooms or other standards for environmental reasons. With such a diverse and complex structure, this policy struggles to be utilized widely. Form-based zoning on the other hand, is new and on the rise. This policy enforces a building's compliance through dimensions and size \cite{zoning}. \\

@@

    \quad Cellular automaton was an idea thrown around to allow zoning cells to connect with one another. The pros of this idea is that it was a way to traverse the grid and set different points to a zone if not in use. As a whole, cellular automata takes in a grid of cells and assigns values, such as on or off to these cells. Starting at a cell, the program finds its neighbors and applies a function to them to determine if their value should be changed. Now, this algorithm is actually very popular for binary systems, such as Conway's Game of Life: seen in Figure \ref{fig:cellular-automaton}. However, it tends to get very complicated very quickly when applied to a system on a higher order, especially, a system such as configurable zoning, which does not have a finite quantity to it. If a user added more zones to the city, the program would struggle to make this functionality possible. Therefore, cellular automata was finally determined not applicable to our zoning scenario.\\

@@

    \quad Unlike zoning, making a network is actually much simpler. The two main steps comprise of creating nodes within the existing zones and making a road system that connects them together. These roads need to ensure that a person can travel from any given node to another. In our case, a node is simply a building. To determine the density of nodes in one district, several values are provided from the user's configurations. These include the average size of one building in the zone. However, this quantity is mainly used to produce the size of the zones. As far as the network goes, the generation of nodes strictly relies on the area of a building in any subzone and the area of a building in the top level zone. If no subzones are present, the types of buildings in the district will be a list of one type (the top level zone). After determining the types of buildings from the configuration file, the program randomly picks a building type and subtracts the area of it from the district's remaining area until no more buildings can be added. However, one important note is that the program also ensures that at least one building of the top level zone's type is present. This is done by scanning neighbors, as seen in section \ref{zoning}. Once the nodes are picked out, they are converted to type Location\footnote{A class defined to make referencing node attributes easier} in the code.\\

@@

    \quad Figure \ref{fig:first-zoning} illustrates the first zoning implementation of the program. This was done by taking the grid and randomly drawing a rectangle by selecting two points. These points were the top left and bottom right corners of the shape. Each rectangle was also constrained to a maximum area. This prevented the rectangle from taking over the entire plot of land. After the rectangle was created, the program selected the largest remaining plot of land. Each plot of land was defined as one edge of the existing rectangle to the end of the city. Repeating this process with the new plot of land allowed for the program to consistently draw out equal sized portions of zones. However, there were several problems with this algorithm. First off, it was overly complex. Looking for available plot of land was computationally intensive and was not always guaranteed to be large enough. In addition, the rectangles did well for three zones, but quickly deteriorated after adding more.

@@

    In this formula, $A$ represents the area of a district, $B$ is the average size of a building in a given zone, and $M$ is the minimum number of buildings within a district. After computing $A$ for every zone, the algorithm picked out the minimum area of a district and set the cell length to \(\sqrt{A}\). This method worked fairly well, but still had several issues. For one, the zones were all of the same size blocks. Also, apart from the predominant zone, all the other districts were usually one block large. In most modern cities, the size of a district depends on its zone, so this was definitely a problem to solve.

@@

    \[C = \frac{A}{m}\]

    In this formula, $C$ represents the maximum number of cells in a given district, $A$ is once again the district area determined above, and $m$ is the minimum building size of all given zones.

@@

    \[N = \frac{A}{B}\]

    In this formula, $N$ is the number of nodes, $A$ is the area of a district\footnote{The district in this iteration was just a rectangle}, and $B$ is the area of one average building in the zone. Thus, each zone had a different density of dots within it.

    \begin{figure}[H]
        \centering
        \vspace{-1em}
        \includegraphics[width=0.44\textwidth]{images/secondzoningwnodes.png}
        \caption{Second iteration of network without roads}
        \label{fig:second-nodes}
    \end{figure}

    \begin{figure}[H]
        \centering
        \vspace{-1em}
        \includegraphics[width=0.44\textwidth]{images/secondzoningwnetwork.png}
        \caption{Second iteration of complete network with roads}
        \label{fig:second-network}
    \end{figure}

    \quad As seen in Figure \ref{fig:second-nodes}, the distribution of buildings stayed relatively similar to the previous iteration. The only difference was in the display of them to the user. The larger dots in the previous iteration made the detail of their location less precise. Thus, a reduction of size was a simple fix to make the data more clear. In Figure \ref{fig:second-network} it can be seen that roads are finally implemented. However, this is not very representative of a real city's road system. The reason for this is because of the way roads are assigned. In each district, one central hub is chosen, and every other node is connected to that hub with a district road. These district roads were not configurable and only allowed cars to travel on them. In addition, the mess of white lines is just a very inefficient highway system. In this case, the central hubs are shuffled and connected to the next hub in a list. Despite the ineffective nature of the road network, it still functioned to complete one of the fundamental goals: having a system that allowed a person to move from one building to another. This data was ported onto MATSIM \cite{matsim}, but it did not reveal anything surprising or interesting. \\

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.44\textwidth]{images/finalzoningwnodes.png}
        \caption{Final iteration of the nodes}
        \label{fig:final-zoning-nodes}
    \end{figure}

@@

    \quad The final iteration of the road system was much neater and more organized with the help of object oriented programming. The nodes were finally calculated with a different algorithm. Given a district, with a size determined in Section \ref{zoning-implementation}, the land was iterated over to find the placement of zones. The program continuously picked random points in the districts and subtracted the area of the placed building from the remaining area in the district. For a zone with subzones within it, the program simply chose the building type from a list of all possible buildings. Once, again all of this information was configurable by the user. In addition, the program placed at least one building from the top level zone in this case by checking neighbors within the district for a building of the correct type. If no building was found, then the program inserted a top level building automatically. This resulted in the variation of dot colors as seen in the plot. The different color represent different building types. Finally, this iteration also added our better road network. This was heavily described in section \ref{network}. However, to put it simply the baseline road network is just a minimum spanning tree that connects nodes in a district. The highways are a minimum spanning tree that connect a number of central hubs in each district, defined in the configuration file. Each type of road also has attributes of speed, capacity, and number of lanes, which can be changed by configurations as well.

@@

    In this equation, $R$ represents the number of additional roads, $r$ is the number of existing roads in a district or highway network, and $M$ is the meshing ratio defined by the configuration file. Once the number of additional roads is determined, the algorithm goes through all edges that have not been added to the minimum spanning tree from the complete graph\footnote{A graph with all possible edges} and randomly chooses roads to add to the network till the $R$ value has been satisfied.

@@

    \quad The final step of the traffic demand generation is creating the people or agents. These agents require schedules to fulfill the requirements of MATSIM \cite{matsim}. Each schedule must be realistic and also highly configurable. To do this, the program takes in inputs of timings of each activity in particular zones. Thus, when nodes are generated and converted to Location objects, as seen in Section \ref{network}, they also are randomly assigned values of times for when agents use the facilities. In addition, the configuration file also determines the number of workers within each zone. Finally, the configuration file also determines the percentage of people that work from within their district and the percentage of people that work from home. These pieces of data are essentially sifted through with if statements to determine the timings of a person's job. However, to first assign an agent a location to work is a little more complicated. First, for each household, all jobs nearby are sorted from closest to furthest. Then, each location is assigned a weight based on its proximity or index. Finally, the algorithm chooses a random job with the weights assigned. A person is determined to work at home by checking if a random quantity in the range \([0, 1)\) is less than the ratio of the population that work from home. If so, the step of picking a job is skipped. Finally, agents are put through an infinite loop, which has a probability of a third to end every iteration. This loop adds leisure activities after work that fit into the schedule of the agent. These activities are chosen the same way as work. However, this time the number of workers working at the facility does not matter. To increase diversity in schedules, the variability of a person arriving to and leaving from a location can also be modified. Since buildings can have multiple types, people can work and leisure at the same location. This allows for a completely customizable schedule of times and places.

@@

    \quad Once the traffic demand has been generated, the information can be utilized to simulate the city. As mentioned prior, MATSIM \cite{matsim} is the agent based model of choice for this program. The data produced is converted into XML format for the simulation. These files are stored within the output directory and can be used by MATSIM later on. In fact, the code even automatically executes the simulation if the executable file is found in the main directory. These results are stored in the new outputs directory and can be analyzed for research. In fact, most of this information can be fed into a wrapper script for implementation of a machine learning algorithm. Such an algorithm can stochastically change the configurations and find the effects on traffic. The Simultaneous Perturbation Stochastic Approximation is an example of such an algorithm to find a more ideal city from a starting slate. However, with such a large set of configuration options, a Simulated Annealing algorithm may work better. Despite this, the data from a random city can be found in the figures below.

@@

    \quad The leg histogram found in Figure \ref{fig:leg-histogram} is one of an inefficient city. Nonetheless, it enunciates the different values and makes it clear the meaning of the colors. The red lines represent agents departing from their source locations for any new task in the schedule. The green lines are the number of agents on route. Finally, the blue lines are the number of agents arriving at their destinations. Since this simulation was of a configuration that restricted the bike's speed to 0.5 meters per second, the people took a long time to reach their destinations. It is also clear that due to the small number of jobs in this city, many agents are leaving at around the same time to get to work, however, the schedules in the evenings get a lot more sparse. \\

@@

    \quad This project was successful in that it captured and modeled a theoretical city from a very easily configurable json file. Of course, there are always things that can be done better. For one, the zoning policy is enforced on pixels that are the size of the largest possible buildings. However, this doesn't work in all cases. Sometimes, housing zones may take up less space than one industrial building. To account for this, a better system of DFS needs to be implemented to ensure that each zone can break down to different size cells. In addition, one of the major flaws of this program is multi-modal transportation. Though, the pipeline does take biking and cars into account, a network of public transport systems would be an amazing update for traffic engineers. The only difficulty with this is allowing the user to easily configure the timings of each public transport line. Despite these setbacks, the project went through several iterations and does a great job of capturing some of the most complex details of a city. Along with fitting in smoothly with MATSIM, the project can be very much added upon with a wrapper stochastic approximation script. This could reveal aspects of city design that may currently be unknown.
